---
title: "Testing New Operators using the MOEADr Package"
author: "Claus Aranha"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing New Operators using the MOEADr Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Let us show how to prepare a new component for the MOEADr framework, 
add it to an existing algorithmic composition, and perform some basic comparisons. 
For this example, we will create a simple variation operator that does not 
exist in the package in its current state, but the same principle works 
for adding components of other classes.

## Creating a new operator

Consider the following "*Gaussian Mutation*" operator: given a set $X$ of
solutions $\mathbf{x_i} \in X$ we add, with probability $p$, a gaussian noise
$r_{ij} \sim \mathcal{N}(\mu, \sigma)$ to each $x_{ij} \in \mathbf{x_i} \in X$. The *R* code 
for this operator is as follows:

```{r}
variation_gaussmut <- function(X, mean = 0, sd = 0.1, p = 0.1, ...) {
  # You want to do some error checking on the parameters here
  # But for the sake of brevity in this case study, we are skipping it.
  
  R <- rnorm(length(X),    # vector of normally distributed values, 
             mean = mean,  # length(R) = nrow(X) * ncol(X)
             sd   = sd)
  R <- R * (runif(length(X)) <= p) # Apply binary mask, probability = p
  return (X + R)                   # Add mutations to the solution matrix
}
```

We would like to highlight a few characteristics of the code sample above.
First, the name of the function must be in the form *variation_[functionname]*.
The MOEADr package uses the function name prefixes to perform some automated
functions such as listing existing components and error checking. The list of
current function prefixes and their meaning is:

- *constraint_*: Constraint handling components
- *decomposition_*: Decomposition functions
- *ls_*: Local search operators
- *scalarization_*: Scalarization functions
- *stop_*: Stop criteria components
- *uptd_*: Update components
- *variation_*: Variation operators

Second, the parameters in the function definition must include: the solution set
matrix $X$, the local parameters for the function, and finally an ellipsis argument to
catch any other inputs passed by the main *moead()* call, such as objective
values and former solution sets. If you want examples of using these parameters
please look at the source code for some of the variation operators included, such as the Binomial Recombination operator (*variation_binrec*)

Other component classes follow similar rules. Please look at existing
implementations as basis for new ideas.

## Testing the new operator

The **MOEADr** package first searches for operators in the base R environment.
Therefore, if you have named your component correctly, all you need to do is add
it to the appropriate parameter in the *moead()* call.

For example, let us replace the variation stack of the original MOEA/D by our Gaussian Mutation operator, followed by simple truncation, and test it on a standard benchmark function (performing a qualitative graphical comparison against the original MOEA/D just for fun):

```{r}
suppressPackageStartupMessages(library(MOEADr))
suppressPackageStartupMessages(library(smoof))

ZDT1 <- make_vectorized_smoof(prob.name  = "ZDT1",
                              dimensions = 30)
problem.zdt1  <- list(name       = "ZDT1",
                      xmin       = rep(0, 30),
                      xmax       = rep(1, 30),
                      m          = 2)


myvar <- list()                                   # Initialize variation stack
myvar[[1]] <- list(name = "gaussmut", p = 0.5)    # Our new operator
myvar[[2]] <- list(name = "truncate")             # Truncation repair operator

results.orig  <- moead(problem  = problem.zdt1,
                       preset   = preset_moead("original"),
                       showpars = list(show.iters = "none"), 
                       seed     = 42)
results.myvar <- moead(problem  = problem.zdt1,
                       preset   = preset_moead("original"),
                       variation = myvar,
                       showpars = list(show.iters = "none"), 
                       seed     = 42)
```

Since the function that implements the Gaussian Mutation operator was defined in
the main environment, and has the required *variation_* prefix, all that we need
to do to use it in the *moead()* function is to add the necessary parameters to
the variation stack. The figures below show the final Pareto front for both the 
standard MOEA/D and the MOEA/D with a Gaussian Mutation operator. From these images, it seems (rather unsurprisingly) that the new operator still needs some work.

```{r, fig.width = 5.5, fig.height = 5.5, fig.align="center"}
plot(results.orig, suppress.pause = TRUE)
plot(results.myvar, suppress.pause = TRUE)
```

We welcome new contributions to **MOEADr**'s component library: users are
invited and encouraged to contact us to add their published contributions to the
package (or simply clone the GitHub repository and submit his or her contribution as a pull request).


